Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators

### 2025-08-01 literature/paper_draft.qmd
refining whole sim section

- Talk about the history and evolution of forecasting hierarchical time series, starting from the early heuristic methods to the modern statistical approaches.

  - Single level methods
  - Optimal combination methods (OLS, WLS)
  - MinT
  - Bayesian, Machine learning
  - Probabilistic methods

@@

To overcome these issues, forecast reconciliation was introduced by @Hyndman2011-jv, and later developed by @Hyndman2016-ic, @Wickramasuriya2019-xq, and ... to achieve coherency in point forecasts and enhance accuracy. Forecast reconciliation projects a collection of independent base forecasts into a set of coherent forecasts that respect the linear constraints defining a hierarchical or grouped time-series system.

@@

- Discuss the interests in MinT and how it has become a standard method for forecast reconciliation.

@@

- Talk about the paper outline and what will be covered in the following sections.

@@

@@

The 100‐series design employs a deliberately intricate aggregation path to to stress‐test reconciliation methods. We first sum the one hundred bottom series into ten intermediate series by grouping them in contiguous blocks of ten. These ten series are then organised into three level-2 aggregates—four, three, and four series, respectively—before finally summing to a single top node. This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both, emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories.

Across all designs, the VAR(1) coefficient matrices remain block-diagonal, ensuring that series within a block interact but blocks remain orthogonal to each other. To save space, we illustrate only the six-by-six's VAR(1) and correlation configuration in @fig-settings-6x6; the two-by-two and the one-hundred-series structures appear in Appendix @sec-sim-supplement.

@@

p3_A <- plot_heatmap(sim3_1$A) +
  ggtitle("50x2 VAR(1) Coefficients Matrix")

p3_Sigma <- plot_heatmap(cov2cor(sim3_1$Sigma), T) +
  ggtitle("50x2 Correlation Matrix")
```

```{r fig settings 6x6}
#| label: fig-settings-6x6
#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 50x2 structure."
#| fig.height: 4
#| fig.width: 6
#| out.width: "85%"
#| fig.align: "center"

@@

For each hierarchy, we simulate two panel lengths, $T=54$ and $T=304$, reserving the final four observations as an out-of-sample test set. In each Monte Carlo replication ($M=500$), we fit univariate ARIMA models (base models) to the training observations using an automatic AICc minimization algorithm from the fabletools package [@O-Hara-Wild2024-we], generating incoherent 1–4-step base forecasts. We then reconcile these forecasts under three covariance estimators: the raw sample covariance (mint_sample), the shrinkage estimator (mint_shr), and the NOVELIST estimator (mint_n).

@@

```{r fig-sim-results-1}
#| label: fig-sim-results-1
#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the 2x2, 6x6, and 50x2 hierarchical structures, for 1 to 4 steps ahead forecasts, with 2 time series lengths (T = 54 and T = 304)."
#| fig.width: 8
#| fig.height: 6
#| out.width: "100%"
#| fig.align: "center"

p_MSE1_1 <- MSE1_1 |> 
  plot_relative_improvement() +
    labs(title = "2x2 Hierarchy", subtitle = "T = 54")
p_MSE2_1 <- MSE2_1 |> 
  plot_relative_improvement() +
    labs(title = "6x6 Hierarchy")
p_MSE3_1 <- MSE3_1 |> 
  plot_relative_improvement() +
    labs(title = "50x2 Hierarchy")

@@

Figure @fig-sim-results-1 illustrates that as hierarchy size and complexity increase, the relative improvements in MSE over the base forecasts from reconciliation amplify, across all 1- to 1-step horizons. While extending the training window (from 50 to 300 in‐sample observations) uniformly improves the accuracy of base ARIMA forecasts, it does not alter the percentage improvements of mint_shr or mint_n. This is evidence that both high-dimensional estimators robustly handle the "large $p \gg T$” regime. By contrast, mint_sample achieves significant accuracy when more data make the sample covariance more reliable.

@@

- 50x2 with usual path and weird path

@@

In our second simulation study, we design a data‐generating process that contrasts “dense” and “sparse” correlation regimes among bottom‐level series, reflecting settings one might encounter in practice. We consider the same hierarchical structure of two large groups of 50 bottom series as above. Specifically, the two groups would have strong within-group dependencies throughout and either modest between‐group correlations (the dense scenario) or complete independence (the sparse scenario). These correlation matrices are depicted in Figure @fig-Sigma-50x2. Both scenarios share the same VAR(1) coefficient structure as in our previous simulations; only the innovation covariance changes. Such a setup mirrors real‐world contexts where, for example, sales within a product line may exhibit strong co‐movements, while those in a separate line operate nearly independently.

@@

```{r}
#| label: fig-Sigma-50x2
#| fig-cap: "Heatmaps of the dense and sparse correlation matrix of the data generating process."
#| fig.height: 12
#| fig.width: 8
#| out.width: "100%"
#| fig.align: "center"
p_dense_Sigma <- sim_dense_50$Sigma |>
  cov2cor() |> 
  plot_heatmap(T) +
    labs(title = "Dense")

@@

```{r}
#| label: fig-sim-results-dsp
#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the 50x2 hierarchical structure with dense and sparse DGP's correlation matrix, for 1 to 4 steps ahead forecasts, with 2 time series lengths (T = 54 and T = 304)."
#| fig.height: 7
#| fig.width: 8
#| out.width: "90%"
#| fig.align: "center"
p_dense_50 <- sim_dense_50$MSE |>
  transform_sim_MSE(F) |> 
  plot_relative_improvement() +
    labs(title = "Dense DGP Covariance Matrix", subtitle = "T = 54")
p_sparse_50 <- sim_sparse_50$MSE |>
  transform_sim_MSE(F) |> 
  plot_relative_improvement() +
    labs(title = "Dense DGP Covariance Matrix")

@@

We also experimented with alternative DGPs: varying group sizes, aggregation paths, and VAR(1) structure. However, to our surprise, most designs failed to distinguish mint_shr from mint_n in any substantive way. Their nearly identical performance under these synthetic scenarios suggests that our current simulation may not unveil the full advantages of the thresholding estimators. This finding motivates our turn to empirical data, where latent structural features and regime shifts, which we will discuss in the next section, may reveal performance differences.

@@

- A table/figure to illustrate the hierarchy

@@

```{r fig settings 50x2}
#| label: fig-settings-50x2
#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 6×6 structure."
#| fig.height: 12
#| fig.width: 8
#| out.width: "85%"
#| fig.align: "center"

>>>>

<!-- - Talk about the history and evolution of forecasting hierarchical time series, starting from the early heuristic methods to the modern statistical approaches.

  - Single level methods
  - Optimal combination methods (OLS, WLS)
  - MinT
  - Bayesian, Machine learning
  - Probabilistic methods -->

@@

To overcome these issues, forecast reconciliation was introduced by @Hyndman2011-jv, and later developed by @Hyndman2016-ic, @Wickramasuriya2019-xq, and others to achieve coherency in point forecasts and enhance accuracy. Forecast reconciliation projects a collection of independent base forecasts into a set of coherent forecasts that respect the linear constraints defining a hierarchical or grouped time-series system.

@@

<!-- - Discuss the interests in MinT and how it has become a standard method for forecast reconciliation.

@@

- Talk about the paper outline and what will be covered in the following sections. -->

@@

For all hierarchies in our experiments, we simulate two panel lengths, $T=54$ and $T=304$, reserving the final four observations as an out-of-sample test set. In each Monte Carlo replication ($M=500$), we fit univariate ARIMA models (base models) to the training observations using an automatic AICc minimization algorithm from the fabletools package [@O-Hara-Wild2024-we], generating incoherent 1–4-step base forecasts. We then reconcile these forecasts under three covariance estimators: the raw sample covariance (mint_sample), the shrinkage estimator (mint_shr), and the NOVELIST estimator (mint_n).

@@

The 100‐series design employs a deliberately intricate aggregation path to stress‐test reconciliation methods. We first sum the one hundred bottom series into ten intermediate series by grouping them in contiguous blocks of ten. These ten series are then organised into three level-2 aggregates—four, three, and four series, respectively—before finally summing to a single top node. This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both, emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories.

To save space, we illustrate only the six-by-six's VAR(1) and correlation configuration in @fig-settings-6x6; the two-by-two and the one-hundred-series structures appear in Appendix @sec-sim-supplement.

@@

p3_A <- plot_heatmap(sim3_1$A) +
  ggtitle("2x50 VAR(1) Coefficients Matrix")

p3_Sigma <- plot_heatmap(cov2cor(sim3_1$Sigma), T) +
  ggtitle("2x50 Correlation Matrix")
```

```{r fig settings 6x6}
#| label: fig-settings-6x6
#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 2x50 structure."
#| fig.height: 4
#| fig.width: 6
#| out.width: "85%"
#| fig.align: "center"

@@

@@

```{r fig-sim-results-1}
#| label: fig-sim-results-1
#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the 2x2, 6x6, and 2x50 hierarchical structures, for 1 to 4 steps ahead forecasts, with 2 time series lengths (T = 54 and T = 304)."
#| fig.width: 8
#| fig.height: 6
#| out.width: "100%"
#| fig.align: "center"

p_MSE1_1 <- MSE1_1 |> 
  plot_relative_improvement() +
    labs(title = "2x2 Hierarchy", subtitle = "T = 54")
p_MSE2_1 <- MSE2_1 |> 
  plot_relative_improvement() +
    labs(title = "6x6 Hierarchy")
p_MSE3_1 <- MSE3_1 |> 
  plot_relative_improvement() +
    labs(title = "2x50 Hierarchy")

@@

Figure @fig-sim-results-1 illustrates that as hierarchy size and complexity increase, the relative improvements in MSE over the base forecasts from reconciliation amplify, across all 1- to 4-step horizons. While extending the training window (from 50 to 300 in‐sample observations) uniformly improves the accuracy of base ARIMA forecasts, it does not alter the relative improvements of mint_shr or mint_n. This is evidence that both high-dimensional estimators robustly handle the "large $p \gg T$” regime. By contrast, mint_sample achieves significant accuracy when more data make the sample covariance more reliable.

@@

- 2x50 with usual path and weird path

@@

In our second simulation study, we design a data‐generating process that contrasts “dense” and “sparse” correlation regimes among bottom‐level series, reflecting settings one might encounter in practice. We consider the same hierarchical structure of two large groups of 50 bottom series as above. Specifically, the two groups would have strong within-group dependencies throughout and either modest between‐group correlations (the dense scenario) or complete independence (the sparse scenario). These correlation matrices are depicted in Figure @fig-Sigma-2x50. Both scenarios share the same VAR(1) coefficient structure as in our previous simulations; only the innovation covariance changes. Such a setup mirrors real‐world contexts where, for example, sales within a product line may exhibit strong co‐movements, while those in a separate line operate nearly independently.

@@

```{r}
#| label: fig-Sigma-2x50
#| fig-cap: "Heatmaps of the dense and sparse correlation matrix of the data generating process."
#| fig.height: 12
#| fig.width: 8
#| out.width: "100%"
#| fig.align: "center"
p_dense_Sigma <- sim_dense_50$Sigma |>
  cov2cor() |> 
  plot_heatmap(T) +
    labs(title = "Dense")

@@

```{r}
#| label: fig-sim-results-dsp
#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the 2x50 hierarchical structure with dense and sparse DGP's correlation matrix, for 1 to 4 steps ahead forecasts, with 2 time series lengths (T = 54 and T = 304)."
#| fig.height: 7
#| fig.width: 8
#| out.width: "90%"
#| fig.align: "center"
p_dense_50 <- sim_dense_50$MSE |>
  transform_sim_MSE(F) |> 
  plot_relative_improvement() +
    labs(title = "Dense DGP Covariance Matrix", subtitle = "T = 54")
p_sparse_50 <- sim_sparse_50$MSE |>
  transform_sim_MSE(F) |> 
  plot_relative_improvement() +
    labs(title = "Dense DGP Covariance Matrix")

@@

Additional designs (varying block sizes, aggregation paths, correlation configurations) also failed to separate NOVELIST from Shrinkage. Their nearly identical performance under these synthetic scenarios suggests that our current simulation may not unveil the full advantages of the thresholding estimators. This finding motivates our turn to empirical data, where latent structural features and regime shifts, which we will discuss in the next section, may reveal performance differences.

@@

<!-- - A table/figure to illustrate the hierarchy -->

@@

```{r fig settings 2x50}
#| label: fig-settings-2x50
#| fig-cap: "Heatmaps of the VAR(1) coefficient matrix and correlation matrix for the 6×6 structure."
#| fig.height: 12
#| fig.width: 8
#| out.width: "85%"
#| fig.align: "center"
